\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-10-14 fully-abstract-comp/popl2019/paper.tex
[WIP] Updated figure 10 formatting
Still working on the format.

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review,anonymous,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

@@

\begin{figure}[htb]
  \centering
  \begin{mathpar}
    \inferrule{ \src{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr)} \\
               \src{\lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta}  \\
               \src{\lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr]} \\
               \src{\perm \in \{\rwx,\rx\}} \\
               \src{\Phi.\mem(\aaddr,\dots,\addr + \calllen-1) = \scall[0]{\offpc,\offsigma}{r_1}{r_2}, \dots, \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}
 }
              { \src{\Phi}  \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}} }
  \end{mathpar}
  \begin{tabular}{|>{$}c<{$}|>{$}p{3.8cm}<{$}|>{\raggedright\arraybackslash}p{6.4cm}|}
    \hline
    i \in \Instr                                 & \sem{i}(\Phi) & Conditions\\
    \hline
    \halt                                        & \halted & \\
    \hline
    \multicolumn{3}{|c|}{\dots\text{ (the operational semantics of \trgcm)}} \\
    \hline
    \store{r_1}{r_2}                             & \srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
    \hline
    \cca{r}{\rn}                                 &\srcalt{\updPcAddr{\Phi\updReg{r}{w}}} &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
    \hline
    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}} & \srcalt{\var{xjmpRes}(c_1,c_2,\Phi')} & $\srcalt{ \Phi' =\var{callConditions}(r_1,r_2,\offpc,\offsigma,\Phi)}$\\
    \hline 
    \multicolumn{3}{|c|}{\dots} \\
    \hline
    \_                                           & \failed & \totherwise \\
    \hline
  \end{tabular}

@@

      \src{r_1 \neq \rtmp{1}}\\
      \src{\reg(r_1) = \sealed{\sigma,c_1} }\\
      \src{r_2 \neq \rtmp{1} }\\
      \src{\reg(r_2) = \sealed{\sigma',c_2} }\\

@@

      \src{\ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
      \src{\ms_{\stk,\var{rest}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} }\\
      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
      \src{c_\var{priv\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\

@@

      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) }\\
      \src{\stk' = (\opc,\ms_{\stk,\priv}) :: \stk }\\

@@

      \src{\sigma_\baddr \leq \sigma \leq \sigma_\eaddr }\\
      \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)}}\\ 
      \src{s_c = \sealed{\sigma,c_\opc} }\\
      \src{s_d = \sealed{\sigma,c_\var{priv\_data}} }\\
      \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}}
    {\src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}}

@@

\begin{multline*}
  \srcxjmpres{c_1,c_2,\Phi} = \\
    \arraycolsep=1.4pt
    \begin{array}{l c >{\raggedright\arraybackslash}p{9.5cm}}
      \arraycolsep=0pt
    \begin{array}{rl}
      \Phi & \updReg{\pcreg}{c_1}\\
           & \updReg{\rdata}{c_2}

@@

    &\phantom{mak} &  $\nonExec{c_2}$ \srcalt{and $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$}\\
      & & \\
    \sourcecolor\left.
      \arraycolsep=0pt
      \begin{array}[t]{rl}
        \Phi' & \updReg{\pcreg}{c_\opc}\\
              & \updReg{\rstk}{c_\stk}\\
              & \updReg{\rdata}{0}\\
              & \updReg{\rtmp{1}}{0}\\
              & \updReg{\rtmp{2}}{0}

@@

    & &
        \srcalt{$c_1 = \retptrc(\baddr,\eaddr,\opc)$ and
        $c_2 = \retptrd(\aaddr_\stk,\eaddr_{\stk})$ and
        $\Phi(\rstk) = \stkptr{\rw,\stkb,\aaddr_\stk-1,\_}$ and
        $\stkb \leq \aaddr_\stk-1$ and
        $\Phi = (\mem,\reg,\stkf::\stk,\ms_\stk)$ and
        $\stkf = (\opc,\ms_{\stk,\priv})$ and
        $c_\opc = ((\rx,\normal),\baddr,\eaddr,\opc)$ and
        $\dom(\ms_{\stk,\priv}) = [\aaddr_\stk,\eaddr_{\stk}]$ and
        $c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk},\aaddr_\stk}$ and
        $\Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk)$ }\\
      & & \\
      \failed &  & \totherwise

@@

\end{multline*}
  \caption{An excerpt of the operational semantics of \srccm{}.}

>>>>

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

@@

\begin{figure}[htb]
  \centering
  \begin{mathpar}
    \inferrule{ \src{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr)} \\
               \src{\lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta}  \\
               \src{\lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr]} \\
               \src{\perm \in \{\rwx,\rx\}} \\
               \src{\Phi.\mem(\aaddr,\dots,\addr + \calllen-1) = \scall[0]{\offpc,\offsigma}{r_1}{r_2}, \dots, \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}
 }
              { \src{\Phi}  \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}} }
  \end{mathpar}
  \begin{tabular}{|>{$}c<{$}|>{$}p{3.8cm}<{$}|>{\raggedright\arraybackslash}p{6.4cm}|}
    \hline
    i \in \Instr                                 & \sem{i}(\Phi) & Conditions\\
    \hline
    \halt                                        & \halted & \\
    \hline
    \multicolumn{3}{|c|}{\dots\text{ (the operational semantics of \trgcm)}} \\
    \hline
    \store{r_1}{r_2}                             & \srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
    \hline
    \cca{r}{\rn}                                 &\srcalt{\updPcAddr{\Phi\updReg{r}{w}}} &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
    \hline
    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}} & \srcalt{\var{xjmpRes}(c_1,c_2,\Phi')} & $\srcalt{ \Phi' =\var{callConditions}(r_1,r_2,\offpc,\offsigma,\Phi)}$\\
    \hline 
    \multicolumn{3}{|c|}{\dots} \\
    \hline
    \_                                           & \failed & \totherwise \\
    \hline
  \end{tabular}
\begin{multline*}
  \srcxjmpres{c_1,c_2,\Phi} = \\
  \left\{ 
    \arraycolsep=1.4pt
    \begin{array}{l c >{\raggedright\arraybackslash}p{9.5cm}}
      \arraycolsep=0pt
    \begin{array}{rl}
      \Phi & \updReg{\pcreg}{c_1}\\
           & \updReg{\rdata}{c_2}
    \end{array}
    &\phantom{mak} &  $\nonExec{c_2}$ \srcalt{and $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$}\\
        \src{\Phi'} 
      & &
        $\src{ \Phi' = \var{returnCondition}(c_1,c_2,\Phi)}$ \\
      \failed &  & \totherwise
    \end{array} 
\right.
\end{multline*}
  \caption{An excerpt of the operational semantics of \srccm{}.}
  \label{fig:source-op-sem}
\end{figure}

\begin{figure}

@@

      \src{\exists \sigma \ldotp \reg(r_1) = \sealed{\sigma, c_1} \wedge \reg(r_2) = \sealed{\sigma, c_2}} \\
      \src{r_1 \neq \rtmp{1}} \\ \src{r_2 \neq \rtmp{2}}
    }{
      \src{\var{argumentsSealedPairs}(\reg,r_1,r_2,c_1,c_2)}
    }

@@

      \src{\ms_{\stk,\var{local}} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
      \src{c_\var{local\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
      \src{\ms_{\stk,\var{unused}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} }\\
      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }
    }{
      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk)}
    }

@@

      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) }
    }{
      \src{ \opc, c_\opc = \var{setupOpc(\reg)}}
    }
      \src{\var{callStackFrame} = (\opc,\ms_{\stk,\priv})}\\
      \src{\stk' = \var{callStackFrame} :: \stk }
    }{
      \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}})}      
    }
      \src{\reg(\pcreg) = ((\_,\_),\_,\_,\aaddr) }\\

@@

      \src{\sigma_\baddr \leq \sigma \leq \sigma_\eaddr }
    }{
      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma)}
    }
      \src{s_d = \sealed{\sigma,c_\var{priv\_data}}}\\
      \src{s_c = \sealed{\sigma,c_\opc}}
    }{
     \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc)}
    }    
    % \inferrule{
    %   % \src{r_1 \neq \rtmp{1}}\\
    %   % \src{\reg(r_1) = \sealed{\sigma',c_1} }\\
    %   % \src{r_2 \neq \rtmp{1} }\\
    %   % \src{\reg(r_2) = \sealed{\sigma',c_2} }\\
    %   % \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} }\\
    %   % \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
    %   % \src{\ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
    %   % \src{\ms_{\stk,\var{rest}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} }\\
    %   % \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
    %   % \src{c_\var{priv\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
    %   % \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) }\\
    %   % \src{\opc = \aaddr + \calllen }\\
    %   % \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) }\\
    %   % \src{\stk' = (\opc,\ms_{\stk,\priv}) :: \stk }\\
    %   % \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} }\\
    %   % \src{\baddr \leq  \aaddr+\offpc \leq \eaddr }\\
    %   % \src{\sigma = \sigma_\aaddr + \offsigma }\\
    %   % \src{\sigma_\baddr \leq \sigma \leq \sigma_\eaddr }\\
    %   \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk)}\\
    %   \src{\opc, c_\opc =\var{setupOpc(\reg)}}\\
    %   \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}})}\\
    %   \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma)}\\
    %   \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc)}\\
    %   \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)}}\\ 
    %   \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}
    % {\src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}}

@@

    \src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})} \quad \mathit{iff}\\
      \quad\quad\left\{ 
      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk) \wedge{}}\\
      \src{\opc, c_\opc =\var{setupOpc(\reg)} \wedge{}}\\
      \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}}) \wedge{}}\\
      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma) \wedge{}}\\
      \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc) \wedge{}}\\
      \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)} \wedge{}}\\ 
      \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}

@@

    \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk} = \var{tokenToStkPtr}(\retptrd(\aaddr_\stk,\eaddr_\stk)}
    \src{((\rx,\normal),\baddr,\eaddr,\opc) = \var{tokenToPc}(\retptrc(\baddr,\eaddr,\opc)}
    \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
    \quad\mathit{iff}\quad
      \left\{ \array{l}
        \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
        \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
        \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
        \src{\dom(\ms_\var{local\_stk} \subseteq [\aaddr_\stk,\eaddr_\stk])}
        \endarray
    \src{c_\stk = \var{reconstructStackPointer}(\reg,c)}
    \quad\mathit{iff}\quad
      \left\{ \array{l}
        \src{c = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge{}} \\
        \src{\reg(\rstk) = \stkptr{\rw, \stkb, \aaddr_\stk-1,\_} \wedge{}} \\
        \src{\stkb \leq \addr_\stk \wedge{}} \\
        \src{c_\stk = \var{tokenToStkPtr(c)} }
        \endarray
    \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))} \quad \mathit{iff}\\
      \quad\quad\left\{ 
      \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
      \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
      \src{c_\opc =  \var{tokenToPc}(c_1)} \\
      \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
      \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}

@@

@@

  }{
d
  }
  \caption{The call condition in the operational semantics of \srccm{}.}
